---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Demonstrating the use of the 'method.stratifiedNNLS' Super Learner method


> A new  method to  estimate the  coefficients for the  super learner  and the
> model  to combine  the  individual  algorithms. A  single  algorithm in  the
> library  ('SL.stratify' in  the example  below) is  used to  data adaptively
> identify subgroups  in the observed  data. The other library  algorithms are
> fit with the  full training data, but  the weights can vary  by subgroup. So
> this   differs  from   the  [subsemble](https://github.com/ledell/subsemble)
> method in  that an overall  predictor is fit and  the weights can  change by
> subgroup, whereas subsemble estimates subset specific predictors.

__Authors:__  Geoffrey Ecoto,  [Antoine Chambaz](https://github.com/achambaz),
[Eric Polley](https://github.com/ecpolley)


## Installation

Copy                                                                       the
[method.stratifiedNNLS.R](https://github.com/ecpolley/SuperLearnerExtra/method/method.stratifiedNNLS.R) 
file                                  from                                 the
[SuperLearnerExtra](https://github.com/ecpolley/SuperLearnerExtra)
repository. Source the file into `R`.

## Example

The code  builds upon  `?SuperLearner`.  We  first load  the namespace  of the
package with  name `SuperLearner` package, attach  it on the search  list, and
source the file into `R`. Then, we simulate a training set and a test set.

```{r}
library(SuperLearner)
source("https://raw.githubusercontent.com/ecpolley/SuperLearnerExtra/master/method/method.stratifiedNNLS.R")
set.seed(23432)

## training set
n <- 500
p <- 50
X <- matrix(rnorm(n*p), nrow = n, ncol = p)
colnames(X) <- paste("X", 1:p, sep="")
X <- data.frame(X)
Y <- X[, 1] + sqrt(abs(X[, 2] * X[, 3])) + X[, 2] - X[, 3] + rnorm(n)

## test set
m <- 1000
newX <- matrix(rnorm(m*p), nrow = m, ncol = p)
colnames(newX) <- paste("X", 1:p, sep = "")
newX <- data.frame(newX)
newY <- newX[, 1] + sqrt(abs(newX[, 2] * newX[, 3])) + newX[, 2] -
  newX[, 3] + rnorm(m)
```

We  define a  library of  algorithms and  run the  super learner  algorithm to
obtain a meta-algorithm that performs almost  as well as the best algorithm in
the library. We rely on the `method.NNLS` method to determine the coefficients
for the super learner and the model to combine the individual algorithms.

```{r}
## generate library and run Super Learner
SL.library <- c("SL.glm", "SL.randomForest", "SL.gam",
                "SL.polymars", "SL.mean")
(test <- SuperLearner(Y = Y, X = X, newX = newX, SL.library = SL.library,
                      verbose = TRUE, method = "method.NNLS"))
```

We  would like  that the  coefficients  for the  super learner  can differ  by
subgroup. The subgroups  are characterized by a  stratification algorithm that
is encoded just  like any other algorithm in the  library, see the `functions`
`SL.stratify`  and `predict.SL.stratify`  below (and  note that  both **must**
output `integer`s). Then,  the new `method` is obtained through  a call to the
`method.stratifiedNNLS` `function`.

```{r}
## CAUTION: 'SL.stratify' and 'predict.SL.stratify' MUST output integers

SL.stratify <- function(Y, X, newX, family, obsWeights, id, ...) {
  if (!is.matrix(X)) {
    X <- as.matrix(X)
  }
  X <- sign(X[, 2] * X[, 3])
  
  if (!is.matrix(newX)) {
    newX <- as.matrix(newX)
  }
  pred <- as.integer(sign(newX[, 2] * newX[, 3]))
  fit <- list(object = X)
  out <- list(pred = pred, fit = fit)
  class(out$fit) <- "SL.statify"
  return(out)
}
environment(SL.stratify) <- asNamespace("SuperLearner")


predict.SL.stratify <- function (object, newdata, family, X = NULL, Y = NULL, ...) {
  if (!is.matrix(newdata)) {
    newdata <- as.matrix(newdata)
  }
  pred <- as.integer(sign(newdata[, 2] * newdata[, 3]))
  return(pred)
}
environment(predict.SL.stratify) <- asNamespace("SuperLearner")

method.stratif.NNLS <- method.stratifiedNNLS("SL.stratify")
environment(method.stratif.NNLS) <- asNamespace("SuperLearner")
```

To compute  the new super  learner, it  suffices to append  the stratification
algorithm to  the library previously  defined, and to  request the use  of the
`method.stratif.NNLS` method.

```{r}

(test.stratified <- SuperLearner(Y = Y, X = X, newX = newX,
                                 SL.library = c(SL.library, "SL.stratify"),
                                 verbose = TRUE,
                                 method = method.stratif.NNLS))
```
The stratum-specific weights differ starkly.

